---
title: "MODULE 1 - STATISTICAL SEGMENTATION"
author: "Melisa Cardenas Pijo"
date: "19/4/2020"
output:
  md_document:
    variant: markdown_github
---
### --- COMPUTING RECENCY, FREQUENCY, MONETARY VALUE ---------
### Loading the dataset

```{r}
data = read.delim(file = 'purchases.txt', header = FALSE, sep = '\t', dec = '.')
```

#### Display the data before transformation
```{r}
head(data)
str(data)
```

### Transforming the dataset
#### Add headers and interpret the last column as a date, extract year of purchase

```{r}
colnames(data) = c('customer_id', 'purchase_amount', 'date_of_purchase')
data$date_of_purchase = as.Date(data$date_of_purchase, "%Y-%m-%d")
data$days_since       = as.numeric(difftime(time1 = "2016-01-01",
                                            time2 = data$date_of_purchase,
                                            units = "days"))
```
Function difftime calculates a difference of two date/time.

### Exploring the transformed dataset
#### Dataset size
```{r}
dim(data)
```

##### Display the data after transformation
```{r}
head(data)
summary(data)
```

### Compute key marketing indicators using SQL language
#### Let's compute key marketing indicators: Recency, Frequency, and Monetary Value (average purchase amount):
```{r}

library(sqldf)

# Compute recency, frequency, and average purchase amount
customers = sqldf("SELECT customer_id,
                          MIN(days_since) AS 'recency',
                          COUNT(*) AS 'frequency',
                          AVG(purchase_amount) AS 'amount'
                   FROM data GROUP BY 1")
head(customers)
```

### Computing key marketing indicators with RSTUDIO
#### Let's compute key marketing indicators: Recency, Frequency, and Monetary Value (average purchase amount):
```{r}
library(dplyr)

customer_id_groups <- group_by(data, customer_id) # first groups data by customer_id

customers_1 <- summarise(customer_id_groups,# then summarizes and creates new variables 
                       recency = min(days_since),     # number of days since last purchase
                       frequency = n(),                # number of purchases
                       amount = mean(purchase_amount))  # average purchase amount (monetary value)

head(customers_1)
```

### Explore the data
```{r}
head(customers)
summary(customers)
hist(customers$recency)
hist(customers$frequency)
hist(customers$amount)
hist(customers$amount, breaks = 100)
```

#### Doing more data exploration
```{r}
library(ggplot2)

ggplot(data=customers, aes(recency)) + geom_histogram(color = I('black'), fill= I('blue3'))
```
```{r}
ggplot(data=customers, aes(frequency)) + geom_histogram(color = I('black'), fill= I('orange'))
```
```{r}
ggplot(data=customers, aes(amount)) + geom_histogram(binwidth = 10, color = I('black'), fill= I('deepskyblue')) +
   scale_x_continuous(limits = c(0, 500),
                     breaks = seq(0, 500, 20))
```
The frequency and monetary value distribution are skewed. It requires data transformation (log).

###--- PREPARING AND TRANSFORMING DATA ----------------------

#### Copying customer data into new data frame

```{r}
new_data = customers
```

#### # Remove customer id as a variable, store it as row names
And storing it as row names:

```{r}
head(new_data)
row.names(new_data) = new_data$customer_id
new_data$customer_id = NULL
head(new_data)

```

#### Take the log-transform of the amount, and plot

```{r}
library(ggplot2)
new_data$amount = log(new_data$amount)
hist(new_data$amount)
 

```
#### Standardize variables

```{r}
new_data = scale(new_data)
head(new_data)
```

### --- RUNNING A HIERARCHICAL SEGMENTATION ------------------

#### Compute distance metrics on standardized data
#### This will likely generate an error on most machines
#### d = dist(new_data)

#### Take a 10% sample

```{r}
sample = seq(1, 18417, by = 10)
head(sample)
customers_sample = customers[sample, ]
new_data_sample  = new_data[sample, ]
```

#### Compute distance metrics on standardized data

```{r}
d = dist(new_data_sample)
```

#### Perform hierarchical clustering on distance metrics
```{r}
c = hclust(d, method="ward.D2")

```

#### Plot de dendogram
```{r}
plot(c)
```

#### Cut at 9 segments

```{r}
members = cutree(c, k = 9)
head(members)
```

#### Show 30 first customers, frequency table
```{r}
members[1:30]
table(members)
```

#### Show profile of each segment
```{r}
aggregate(customers_sample[, 2:4], by = list(members), mean)
```

Groups by segment number the variables (recency, frequency and monetary value) of the non-normalized customer sample and shows the mean value of each variable for each segment

### Analysis

The above table summarizes the diversity of profile existing between the 18,417 customers of the database. For example:

*Segment 4 has an average recency of 162 days, an average of 2.4 purchases made in the past and an average purchase amount of $41.10

*Segment #6 spends much more in the store whereas segment #2 spends much less.

*Segment #3 made a huge number of purchase in the past whereas segment #8 made only one, and a long time ago.











